### 2019-10-29 thesis/content/abstract.tex
added thesis wireframe

% !TEX root = ../main.tex
%
\pdfbookmark[0]{Abstract}{Abstract}
\chapter*{Abstract}%
\label{sec:abstract}
\vspace*{-10mm}