### 2019-10-29 thesis/content/chapter-appendix.tex

% !TEX root = ../main.tex
% chktex-file 46

\chapter{Appendix}%
\label{sec:appendix}