### 2019-10-29 thesis/content/chapter-conclusion.tex

% !TEX root = ../main.tex
% chktex-file 46
\chapter{Conclusion}%
\label{sec:conclusion}

\section{Review}%
\label{sec:conclusion:review}

\section{Future Directions}%
\label{sec:conclusion:todo}