### 2019-10-29 thesis/main.tex

% **************************************************
% Document Class Definition
% **************************************************
\documentclass[%
    paper=A4,               % paper size --> A4 is default in Germany
    twoside=true,           % onesite or twoside printing
    openright,              % doublepage cleaning ends up right side
    parskip=full,           % spacing value / method for paragraphs
    chapterprefix=true,     % prefix for chapter marks
    11pt,                   % font size
    headings=normal,        % size of headings
    bibliography=totoc,     % include bib in toc
    listof=totoc,           % include listof entries in toc
    titlepage=on,           % own page for each title page
    captions=tableabove,    % display table captions above the float env
    draft=false,            % value for draft version
]{scrreprt}%

\input{config}

% **************************************************
% Document CONTENT
% **************************************************
\begin{document}

% --------------------------
% rename document parts
% --------------------------

\newcommand{\fixedinput}{\input}

% --------------------------
% Front matter
% --------------------------
\pagenumbering{roman}			% roman page numbing (invisible for empty page style)
\pagestyle{empty}				% no header or footers
\input{content/titlepages}		% INCLUDE: all titlepages
\cleardoublepage\

\pagestyle{plain}				% display just page numbers
\input{content/abstract}		% INCLUDE: the abstracts (english and german)
\cleardoublepage\
%
\setcounter{tocdepth}{2}		% define depth of toc
\tableofcontents				% display table of contents
\cleardoublepage\
% --------------------------
% Body matter
% --------------------------
\pagestyle{maincontentstyle} 	% fancy header and footer

\input{content/chapter-introduction}
\input{content/chapter-conclusion}

% --------------------------
% Back matter
% --------------------------
\appendix\cleardoublepage\
\input{content/chapter-appendix}
%
{%
\setstretch{1.1}
\renewcommand{\bibfont}{\normalfont\small}
\setlength{\biblabelsep}{0pt}
\setlength{\bibitemsep}{0.5\baselineskip plus 0.5\baselineskip} % chktex 1
\setcounter{biburllcpenalty}{9000}
\setcounter{biburlucpenalty}{9999}
\printbibliography[nottype=online]
\printbibliography[heading=subbibliography,title={Webseiten},type=online]
}
\cleardoublepage\

\listoffigures
\cleardoublepage\

\input{content/declaration}

% **************************************************
% End of Document CONTENT
% **************************************************
\end{document}