### 2020-03-09 thesis/content/abstract.tex
started with ltag definition + various minor fixes

%!TEX root = ../main.tex
%
\pdfbookmark[0]{Abstract}{Abstract}
\chapter*{Abstract}%
\label{sec:abstract}
\vspace*{-10mm}