### 2020-03-09 thesis/content/chapter-appendix.tex

%!TEX root = ../main.tex
% chktex-file 46

\chapter{Appendix}%
\label{sec:appendix}