### 2020-04-07 thesis/content/chapter-appendix.tex
started with synthetic evaluation section + more kernel evals

%!TEX root = ../main.tex
% chktex-file 46

\chapter{Appendix}%
\label{sec:appendix}

\section{Dataset Statistics and Descriptions}%
\label{sec:appendix:ds-stats}

\begin{table}[ht]
	\caption{Sizes of the evaluated binary classification datasets and their graphs.}\label{tbl:appendix:ds-stats}
	\csvreader[
		tabular={lrrrrrrrr},
		separator=comma,
		table head={%
			\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & &\multicolumn{3}{c}{vertex count $\left|\mathcal{V}_G\right|$} & \multicolumn{3}{c}{edge count $\left|\mathcal{E}_G\right|$} \\%
			& \multirow{-2}{*}[-0.2em]{\shortstack[r]{no.\ of\\ graphs}} & \multirow{-2}{*}[-0.2em]{\shortstack[r]{no.\ of vertex\\ feature dims.}} & $\min$ & $\mean$ & $\max$ & $\min$ & $\mean$ & $\max$ \\\toprule%
		},
		table foot=\bottomrule,
		late after line=\\
	]{data/ds_stats.csv}%
	{name=\name,graph_count=\gcount,%
	node_count_min=\ncountmin,node_count_mean=\ncountmean,node_count_max=\ncountmax,%
	edge_count_min=\ecountmin,edge_count_mean=\ecountmean,edge_count_max=\ecountmax,%
	node_degree_min=\ndegmin,node_degree_mean=\ndegmean,node_degree_max=\ndegmax,%
	dim_node_features=\nfdim,dim_edge_features=\efdim%
	}%
	{\textbf{\name}&\gcount&%
	\nfdim&%
	\ncountmin&\ncountmean&\ncountmax&%
	\ecountmin&\ecountmean&\ecountmax%
	}
\end{table}

\paragraph{TRIANGLE}
The triangle detection dataset was generated by sampling three graphs with exactly one unicolored triangle uniformly at random for each possible combination of the following parameters:
The number of vertices (between 6 and 32), the vertex color proportions (either 50/50\%, 75/25\% or 25/75\% vertices with the color \colorlabel{t_blue}{A}/\colorlabel{t_red}{B}), the graph density (${\left| \mathcal{V}_{G} \right|}^{-2} \left| \mathcal{E}_{G} \right| \in \{ \nicefrac{1}{4}, \nicefrac{1}{2} \}$) the graph class (add a triangle with either the color \colorlabel{t_blue}{A} or \colorlabel{t_red}{B}).