### 2020-04-10 thesis/data/diffs/imdb.tex

% This file was generated by the LTAG results postprocessor. Do not edit manually.
{\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
\begin{tabular}{lccccccccccccc}
& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.7\\ \pm 3.2}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-1.2\\ \pm 2.6}} & {\tiny\Vectorstack{+0.6\\ \pm 3.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 4.6}}} & {\tiny\Vectorstack{+4.2\\ \pm 6.1}} & {\tiny\Vectorstack{-0.4\\ \pm 3.1}} & {\tiny\Vectorstack{+0.1\\ \pm 2.4}} & {\tiny\Vectorstack{+0.1\\ \pm 3.5}} & {\tiny\Vectorstack{-0.2\\ \pm 4.4}} & {\tiny\Vectorstack{-1.2\\ \pm 3.1}} & {\tiny\Vectorstack{-0.1\\ \pm 3.8}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&{\tiny\Vectorstack{+1.7\\ \pm 3.2}} &  & {\tiny\Vectorstack{-0.9\\ \pm 3.7}} & {\tiny\Vectorstack{+0.5\\ \pm 3.3}} & {\tiny\Vectorstack{+2.3\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 4.6}}} & {\tiny\Vectorstack{+5.9\\ \pm 6.1}} & {\tiny\Vectorstack{+1.3\\ \pm 3.8}} & {\tiny\Vectorstack{+1.8\\ \pm 4.3}} & {\tiny\Vectorstack{+1.8\\ \pm 4.1}} & {\tiny\Vectorstack{+1.5\\ \pm 4.6}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & {\tiny\Vectorstack{+1.6\\ \pm 5.1}} \\[2pt]
\textbf{WL\textsubscript{SP}} ($T=5$)&{\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+0.9\\ \pm 3.7}} &  & {\tiny\Vectorstack{+1.4\\ \pm 4.9}} & {\tiny\Vectorstack{+3.2\\ \pm 5.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 5.0}}} & {\tiny\Vectorstack{+6.8\\ \pm 6.5}} & {\tiny\Vectorstack{+2.2\\ \pm 4.8}} & {\tiny\Vectorstack{+2.7\\ \pm 4.4}} & {\tiny\Vectorstack{+2.7\\ \pm 5.0}} & {\tiny\Vectorstack{+2.4\\ \pm 5.7}} & {\tiny\Vectorstack{+1.4\\ \pm 5.1}} & {\tiny\Vectorstack{+2.5\\ \pm 5.6}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+1.2\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 3.3}} & {\tiny\Vectorstack{-1.4\\ \pm 4.9}} &  & {\tiny\Vectorstack{+1.8\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 5.7}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.2}} & {\tiny\Vectorstack{+0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.0\\ \pm 3.0}} & {\tiny\Vectorstack{+1.1\\ \pm 3.7}} \\[2pt]
\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.6\\ \pm 3.5}} & {\tiny\Vectorstack{-2.3\\ \pm 3.2}} & {\tiny\Vectorstack{-3.2\\ \pm 5.0}} & {\tiny\Vectorstack{-1.8\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+3.6\\ \pm 7.1}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-0.5\\ \pm 3.3}} & {\tiny\Vectorstack{-0.8\\ \pm 3.1}} & {\tiny\Vectorstack{-1.8\\ \pm 3.0}} & {\tiny\Vectorstack{-0.7\\ \pm 4.3}} \\[2pt]
\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 4.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-22\\ \pm 4.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 5.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 6.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.5}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 6.9}}} \\[2pt]
\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 6.1}} & {\tiny\Vectorstack{-5.9\\ \pm 6.1}} & {\tiny\Vectorstack{-6.8\\ \pm 6.5}} & {\tiny\Vectorstack{-5.4\\ \pm 7.2}} & {\tiny\Vectorstack{-3.6\\ \pm 7.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 6.3}}} &  & {\tiny\Vectorstack{-4.6\\ \pm 7.5}} & {\tiny\Vectorstack{-4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-4.1\\ \pm 7.3}} & {\tiny\Vectorstack{-4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-5.4\\ \pm 7.0}} & {\tiny\Vectorstack{-4.3\\ \pm 8.4}} \\[2pt]
\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+0.4\\ \pm 3.1}} & {\tiny\Vectorstack{-1.3\\ \pm 3.8}} & {\tiny\Vectorstack{-2.2\\ \pm 4.8}} & {\tiny\Vectorstack{-0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.0}}} & {\tiny\Vectorstack{+4.6\\ \pm 7.5}} &  & {\tiny\Vectorstack{+0.5\\ \pm 2.8}} & {\tiny\Vectorstack{+0.5\\ \pm 2.5}} & {\tiny\Vectorstack{+0.2\\ \pm 2.8}} & {\tiny\Vectorstack{-0.8\\ \pm 2.6}} & {\tiny\Vectorstack{+0.2\\ \pm 2.4}} \\[2pt]
\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-0.1\\ \pm 2.4}} & {\tiny\Vectorstack{-1.8\\ \pm 4.3}} & {\tiny\Vectorstack{-2.7\\ \pm 4.4}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-0.5\\ \pm 2.8}} &  & {\tiny\Vectorstack{+0.0\\ \pm 2.8}} & {\tiny\Vectorstack{-0.3\\ \pm 3.5}} & {\tiny\Vectorstack{-1.3\\ \pm 2.5}} & {\tiny\Vectorstack{-0.2\\ \pm 2.7}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-0.1\\ \pm 3.5}} & {\tiny\Vectorstack{-1.8\\ \pm 4.1}} & {\tiny\Vectorstack{-2.7\\ \pm 5.0}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 3.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.8}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.3}} & {\tiny\Vectorstack{-0.5\\ \pm 2.5}} & {\tiny\Vectorstack{-0.0\\ \pm 2.8}} &  & {\tiny\Vectorstack{-0.3\\ \pm 1.7}} & {\tiny\Vectorstack{-1.3\\ \pm 1.1}} & {\tiny\Vectorstack{-0.3\\ \pm 2.2}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+0.2\\ \pm 4.4}} & {\tiny\Vectorstack{-1.5\\ \pm 4.6}} & {\tiny\Vectorstack{-2.4\\ \pm 5.7}} & {\tiny\Vectorstack{-1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.8\\ \pm 3.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.4}}} & {\tiny\Vectorstack{+4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.3\\ \pm 3.5}} & {\tiny\Vectorstack{+0.3\\ \pm 1.7}} &  & {\tiny\Vectorstack{-1.0\\ \pm 2.0}} & {\tiny\Vectorstack{+0.1\\ \pm 3.4}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+1.2\\ \pm 3.1}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-1.4\\ \pm 5.1}} & {\tiny\Vectorstack{-0.0\\ \pm 3.0}} & {\tiny\Vectorstack{+1.8\\ \pm 3.0}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 5.5}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.0}} & {\tiny\Vectorstack{+0.8\\ \pm 2.6}} & {\tiny\Vectorstack{+1.3\\ \pm 2.5}} & {\tiny\Vectorstack{+1.3\\ \pm 1.1}} & {\tiny\Vectorstack{+1.0\\ \pm 2.0}} &  & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.1\\ \pm 3.8}} & {\tiny\Vectorstack{-1.6\\ \pm 5.1}} & {\tiny\Vectorstack{-2.5\\ \pm 5.6}} & {\tiny\Vectorstack{-1.1\\ \pm 3.7}} & {\tiny\Vectorstack{+0.7\\ \pm 4.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 6.9}}} & {\tiny\Vectorstack{+4.3\\ \pm 8.4}} & {\tiny\Vectorstack{-0.2\\ \pm 2.4}} & {\tiny\Vectorstack{+0.2\\ \pm 2.7}} & {\tiny\Vectorstack{+0.3\\ \pm 2.2}} & {\tiny\Vectorstack{-0.1\\ \pm 3.4}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} & 
\end{tabular}}