### 2020-04-10 thesis/data/diffs/nci.tex

% This file was generated by the LTAG results postprocessor. Do not edit manually.
{\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
\begin{tabular}{lccccccccccc}
& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} \\
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 1.3}}} & {\tiny\Vectorstack{-2.8\\ \pm 1.5}} & {\tiny\Vectorstack{+2.3\\ \pm 2.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-1.9\\ \pm 2.9}} & {\tiny\Vectorstack{-4.4\\ \pm 2.3}} & {\tiny\Vectorstack{+3.1\\ \pm 2.5}} & {\tiny\Vectorstack{+1.6\\ \pm 2.6}} & {\tiny\Vectorstack{+3.7\\ \pm 3.6}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 1.3}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.7\\ \pm 1.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 1.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+18\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.0\\ \pm 2.7}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.6\\ \pm 2.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.1\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 2.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+15\\ \pm 3.8}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{+2.8\\ \pm 1.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.7\\ \pm 1.7}}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+5.1\\ \pm 2.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+0.9\\ \pm 2.0}} & {\tiny\Vectorstack{-1.6\\ \pm 2.3}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+6.5\\ \pm 4.2}} \\[2pt]
\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.3\\ \pm 2.4}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 1.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-5.1\\ \pm 2.3}}} &  & {\tiny\Vectorstack{+3.9\\ \pm 4.1}} & {\tiny\Vectorstack{-5.8\\ \pm 3.6}} & {\tiny\Vectorstack{-4.3\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.7\\ \pm 2.5}}} & {\tiny\Vectorstack{+0.8\\ \pm 3.7}} & {\tiny\Vectorstack{-0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+1.4\\ \pm 4.0}} \\[2pt]
\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-18\\ \pm 2.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.0\\ \pm 3.0}}} & {\tiny\Vectorstack{-3.9\\ \pm 4.1}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.8\\ \pm 2.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.2\\ \pm 2.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.6}}} & {\tiny\Vectorstack{-3.1\\ \pm 3.0}} & {\tiny\Vectorstack{-4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-2.6\\ \pm 5.0}} \\[2pt]
\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+3.5\\ \pm 2.2}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-8.0\\ \pm 2.7}}} & {\tiny\Vectorstack{+0.7\\ \pm 2.7}} & {\tiny\Vectorstack{+5.8\\ \pm 3.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+9.8\\ \pm 2.4}}} &  & {\tiny\Vectorstack{+1.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.9\\ \pm 2.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{+5.1\\ \pm 3.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.2\\ \pm 3.3}}} \\[2pt]
\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+1.9\\ \pm 2.9}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-9.6\\ \pm 2.2}}} & {\tiny\Vectorstack{-0.9\\ \pm 2.0}} & {\tiny\Vectorstack{+4.3\\ \pm 2.5}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+8.2\\ \pm 2.9}}} & {\tiny\Vectorstack{-1.6\\ \pm 3.5}} &  & {\tiny\Vectorstack{-2.4\\ \pm 2.6}} & {\tiny\Vectorstack{+5.1\\ \pm 3.5}} & {\tiny\Vectorstack{+3.5\\ \pm 2.2}} & {\tiny\Vectorstack{+5.6\\ \pm 4.8}} \\[2pt]
\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+4.4\\ \pm 2.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.1\\ \pm 2.0}}} & {\tiny\Vectorstack{+1.6\\ \pm 2.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+6.7\\ \pm 2.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.6}}} & {\tiny\Vectorstack{+0.9\\ \pm 2.2}} & {\tiny\Vectorstack{+2.4\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+7.5\\ \pm 3.1}}} & {\tiny\Vectorstack{+5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+8.1\\ \pm 4.2}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-3.1\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 2.3}}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{-0.8\\ \pm 3.7}} & {\tiny\Vectorstack{+3.1\\ \pm 3.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-6.6\\ \pm 2.5}}} & {\tiny\Vectorstack{-5.1\\ \pm 3.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.5\\ \pm 3.1}}} &  & {\tiny\Vectorstack{-1.5\\ \pm 2.2}} & {\tiny\Vectorstack{+0.6\\ \pm 2.9}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{-1.6\\ \pm 2.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 2.0}}} & {\tiny\Vectorstack{-4.4\\ \pm 2.5}} & {\tiny\Vectorstack{+0.8\\ \pm 2.7}} & {\tiny\Vectorstack{+4.7\\ \pm 3.1}} & {\tiny\Vectorstack{-5.1\\ \pm 3.2}} & {\tiny\Vectorstack{-3.5\\ \pm 2.2}} & {\tiny\Vectorstack{-5.9\\ \pm 3.2}} & {\tiny\Vectorstack{+1.5\\ \pm 2.2}} &  & {\tiny\Vectorstack{+2.1\\ \pm 3.4}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{-3.7\\ \pm 3.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-15\\ \pm 3.8}}} & {\tiny\Vectorstack{-6.5\\ \pm 4.2}} & {\tiny\Vectorstack{-1.4\\ \pm 4.0}} & {\tiny\Vectorstack{+2.6\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-7.2\\ \pm 3.3}}} & {\tiny\Vectorstack{-5.6\\ \pm 4.8}} & {\tiny\Vectorstack{-8.1\\ \pm 4.2}} & {\tiny\Vectorstack{-0.6\\ \pm 2.9}} & {\tiny\Vectorstack{-2.1\\ \pm 3.4}} & 
\end{tabular}}