### 2020-04-10 thesis/data/diffs/triangle.tex

% This file was generated by the LTAG results postprocessor. Do not edit manually.
{\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
\begin{tabular}{lccccccccccccc}
& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=5$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{+2.2\\ \pm 9.2}} & {\tiny\Vectorstack{-4.9\\ \pm 12}} & {\tiny\Vectorstack{+8.2\\ \pm 15}} & {\tiny\Vectorstack{+2.9\\ \pm 12}} & {\tiny\Vectorstack{+20\\ \pm 19}} & {\tiny\Vectorstack{-5.2\\ \pm 16}} & {\tiny\Vectorstack{-12\\ \pm 18}} & {\tiny\Vectorstack{-17\\ \pm 19}} & {\tiny\Vectorstack{-26\\ \pm 15}} & {\tiny\Vectorstack{-28\\ \pm 17}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-26\\ \pm 13}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-35\\ \pm 13}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=5$)&{\tiny\Vectorstack{-2.2\\ \pm 9.2}} &  & {\tiny\Vectorstack{-7.1\\ \pm 14}} & {\tiny\Vectorstack{+6.1\\ \pm 9.8}} & {\tiny\Vectorstack{+0.8\\ \pm 13}} & {\tiny\Vectorstack{+18\\ \pm 14}} & {\tiny\Vectorstack{-7.4\\ \pm 11}} & {\tiny\Vectorstack{-14\\ \pm 16}} & {\tiny\Vectorstack{-19\\ \pm 16}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-28\\ \pm 14}}} & {\tiny\Vectorstack{-30\\ \pm 17}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-28\\ \pm 9.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-37\\ \pm 11}}} \\[2pt]
\textbf{WL\textsubscript{SP}} ($T=5$)&{\tiny\Vectorstack{+4.9\\ \pm 12}} & {\tiny\Vectorstack{+7.1\\ \pm 14}} &  & {\tiny\Vectorstack{+13\\ \pm 13}} & {\tiny\Vectorstack{+7.9\\ \pm 10}} & {\tiny\Vectorstack{+25\\ \pm 16}} & {\tiny\Vectorstack{-0.3\\ \pm 13}} & {\tiny\Vectorstack{-7.1\\ \pm 10}} & {\tiny\Vectorstack{-12\\ \pm 12}} & {\tiny\Vectorstack{-21\\ \pm 11}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 11}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-30\\ \pm 8.9}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-8.2\\ \pm 15}} & {\tiny\Vectorstack{-6.1\\ \pm 9.8}} & {\tiny\Vectorstack{-13\\ \pm 13}} &  & {\tiny\Vectorstack{-5.3\\ \pm 11}} & {\tiny\Vectorstack{+12\\ \pm 7.5}} & {\tiny\Vectorstack{-13\\ \pm 6.7}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 9.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 8.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-36\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-34\\ \pm 7.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-43\\ \pm 6.8}}} \\[2pt]
\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.9\\ \pm 12}} & {\tiny\Vectorstack{-0.8\\ \pm 13}} & {\tiny\Vectorstack{-7.9\\ \pm 10}} & {\tiny\Vectorstack{+5.3\\ \pm 11}} &  & {\tiny\Vectorstack{+17\\ \pm 15}} & {\tiny\Vectorstack{-8.2\\ \pm 13}} & {\tiny\Vectorstack{-15\\ \pm 14}} & {\tiny\Vectorstack{-20\\ \pm 13}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 8.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-31\\ \pm 9.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 10}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-38\\ \pm 9.0}}} \\[2pt]
\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-20\\ \pm 19}} & {\tiny\Vectorstack{-18\\ \pm 14}} & {\tiny\Vectorstack{-25\\ \pm 16}} & {\tiny\Vectorstack{-12\\ \pm 7.5}} & {\tiny\Vectorstack{-17\\ \pm 15}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-25\\ \pm 8.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-32\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-37\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-46\\ \pm 9.8}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-48\\ \pm 14}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-46\\ \pm 9.2}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-55\\ \pm 8.5}}} \\[2pt]
\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+5.2\\ \pm 16}} & {\tiny\Vectorstack{+7.4\\ \pm 11}} & {\tiny\Vectorstack{+0.3\\ \pm 13}} & {\tiny\Vectorstack{+13\\ \pm 6.7}} & {\tiny\Vectorstack{+8.2\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 8.4}}} &  & {\tiny\Vectorstack{-6.8\\ \pm 13}} & {\tiny\Vectorstack{-12\\ \pm 9.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 9.9}}} & {\tiny\Vectorstack{-23\\ \pm 14}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 8.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-29\\ \pm 7.3}}} \\[2pt]
\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{+12\\ \pm 18}} & {\tiny\Vectorstack{+14\\ \pm 16}} & {\tiny\Vectorstack{+7.1\\ \pm 10}} & {\tiny\Vectorstack{+20\\ \pm 13}} & {\tiny\Vectorstack{+15\\ \pm 14}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+32\\ \pm 14}}} & {\tiny\Vectorstack{+6.8\\ \pm 13}} &  & {\tiny\Vectorstack{-5.0\\ \pm 6.6}} & {\tiny\Vectorstack{-14\\ \pm 9.6}} & {\tiny\Vectorstack{-16\\ \pm 11}} & {\tiny\Vectorstack{-14\\ \pm 8.8}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-23\\ \pm 9.8}}} \\[2pt]
\textbf{2-GNN} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+17\\ \pm 19}} & {\tiny\Vectorstack{+19\\ \pm 16}} & {\tiny\Vectorstack{+12\\ \pm 12}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+25\\ \pm 9.4}}} & {\tiny\Vectorstack{+20\\ \pm 13}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+37\\ \pm 9.2}}} & {\tiny\Vectorstack{+12\\ \pm 9.5}} & {\tiny\Vectorstack{+5.0\\ \pm 6.6}} &  & {\tiny\Vectorstack{-8.9\\ \pm 6.3}} & {\tiny\Vectorstack{-11\\ \pm 8.8}} & {\tiny\Vectorstack{-9.2\\ \pm 7.1}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-18\\ \pm 7.2}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{+26\\ \pm 15}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+28\\ \pm 14}}} & {\tiny\Vectorstack{+21\\ \pm 11}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 8.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 8.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+46\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 9.9}}} & {\tiny\Vectorstack{+14\\ \pm 9.6}} & {\tiny\Vectorstack{+8.9\\ \pm 6.3}} &  & {\tiny\Vectorstack{-2.2\\ \pm 7.2}} & {\tiny\Vectorstack{-0.3\\ \pm 6.0}} & {\tiny\Vectorstack{-8.7\\ \pm 4.6}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+28\\ \pm 17}} & {\tiny\Vectorstack{+30\\ \pm 17}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+36\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+31\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+48\\ \pm 14}}} & {\tiny\Vectorstack{+23\\ \pm 14}} & {\tiny\Vectorstack{+16\\ \pm 11}} & {\tiny\Vectorstack{+11\\ \pm 8.8}} & {\tiny\Vectorstack{+2.2\\ \pm 7.2}} &  & {\tiny\Vectorstack{+1.9\\ \pm 9.7}} & {\tiny\Vectorstack{-6.5\\ \pm 8.7}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+26\\ \pm 13}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+28\\ \pm 9.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+34\\ \pm 7.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 10}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+46\\ \pm 9.2}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 8.3}}} & {\tiny\Vectorstack{+14\\ \pm 8.8}} & {\tiny\Vectorstack{+9.2\\ \pm 7.1}} & {\tiny\Vectorstack{+0.3\\ \pm 6.0}} & {\tiny\Vectorstack{-1.9\\ \pm 9.7}} &  & {\tiny\Vectorstack{-8.5\\ \pm 4.7}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+35\\ \pm 13}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+37\\ \pm 11}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+30\\ \pm 8.9}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+43\\ \pm 6.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+38\\ \pm 9.0}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+55\\ \pm 8.5}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+29\\ \pm 7.3}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+23\\ \pm 9.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+18\\ \pm 7.2}}} & {\tiny\Vectorstack{+8.7\\ \pm 4.6}} & {\tiny\Vectorstack{+6.5\\ \pm 8.7}} & {\tiny\Vectorstack{+8.5\\ \pm 4.7}} & 
\end{tabular}}