### 2020-04-11 thesis/data/diffs/reddit.tex

% This file was generated by the LTAG results postprocessor. Do not edit manually.
{\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
\begin{tabular}{lcccccccccc}
& \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)} \\
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=1$)& & {\tiny\Vectorstack{-1.8\\ \pm 1.8}} & {\tiny\Vectorstack{+0.5\\ \pm 4.9}} & {\tiny\Vectorstack{+0.8\\ \pm 3.6}} & {\tiny\Vectorstack{+4.2\\ \pm 10}} & {\tiny\Vectorstack{-11\\ \pm 6.9}} & {\tiny\Vectorstack{+2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-7.4\\ \pm 6.2}} & {\tiny\Vectorstack{-0.9\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.6}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{WL\textsubscript{ST}*}} ($T=3$)&{\tiny\Vectorstack{+1.8\\ \pm 1.8}} &  & {\tiny\Vectorstack{+2.3\\ \pm 4.7}} & {\tiny\Vectorstack{+2.6\\ \pm 3.5}} & {\tiny\Vectorstack{+5.9\\ \pm 11}} & {\tiny\Vectorstack{-9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+0.9\\ \pm 1.6}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-11\\ \pm 2.8}}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-LWL*}} ($T=3$)&{\tiny\Vectorstack{-0.5\\ \pm 4.9}} & {\tiny\Vectorstack{-2.3\\ \pm 4.7}} &  & {\tiny\Vectorstack{+0.3\\ \pm 4.2}} & {\tiny\Vectorstack{+3.7\\ \pm 11}} & {\tiny\Vectorstack{-11\\ \pm 7.3}} & {\tiny\Vectorstack{+2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-7.9\\ \pm 5.5}} & {\tiny\Vectorstack{-1.4\\ \pm 5.0}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-13\\ \pm 3.6}}} \\[2pt]
\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-0.8\\ \pm 3.6}} & {\tiny\Vectorstack{-2.6\\ \pm 3.5}} & {\tiny\Vectorstack{-0.3\\ \pm 4.2}} &  & {\tiny\Vectorstack{+3.3\\ \pm 11}} & {\tiny\Vectorstack{-12\\ \pm 7.6}} & {\tiny\Vectorstack{+2.1\\ \pm 2.4}} & {\tiny\Vectorstack{-8.2\\ \pm 6.2}} & {\tiny\Vectorstack{-1.7\\ \pm 3.3}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-14\\ \pm 3.1}}} \\[2pt]
\textbf{Baseline} ($\mathrm{sum}$)&{\tiny\Vectorstack{-4.2\\ \pm 10}} & {\tiny\Vectorstack{-5.9\\ \pm 11}} & {\tiny\Vectorstack{-3.7\\ \pm 11}} & {\tiny\Vectorstack{-3.3\\ \pm 11}} &  & {\tiny\Vectorstack{-15\\ \pm 12}} & {\tiny\Vectorstack{-1.2\\ \pm 12}} & {\tiny\Vectorstack{-12\\ \pm 13}} & {\tiny\Vectorstack{-5.0\\ \pm 11}} & {\tiny\Vectorstack{-17\\ \pm 11}} \\[2pt]
\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{+11\\ \pm 6.9}} & {\tiny\Vectorstack{+9.0\\ \pm 7.1}} & {\tiny\Vectorstack{+11\\ \pm 7.3}} & {\tiny\Vectorstack{+12\\ \pm 7.6}} & {\tiny\Vectorstack{+15\\ \pm 12}} &  & {\tiny\Vectorstack{+14\\ \pm 8.3}} & {\tiny\Vectorstack{+3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+9.9\\ \pm 7.4}} & {\tiny\Vectorstack{-2.2\\ \pm 7.2}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mean$)&{\tiny\Vectorstack{-2.9\\ \pm 3.5}} & {\tiny\Vectorstack{-4.7\\ \pm 3.3}} & {\tiny\Vectorstack{-2.4\\ \pm 5.5}} & {\tiny\Vectorstack{-2.1\\ \pm 2.4}} & {\tiny\Vectorstack{+1.2\\ \pm 12}} & {\tiny\Vectorstack{-14\\ \pm 8.3}} &  & {\tiny\Vectorstack{-10\\ \pm 7.1}} & {\tiny\Vectorstack{-3.8\\ \pm 2.5}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 3.3}}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mean$)&{\tiny\Vectorstack{+7.4\\ \pm 6.2}} & {\tiny\Vectorstack{+5.7\\ \pm 6.8}} & {\tiny\Vectorstack{+7.9\\ \pm 5.5}} & {\tiny\Vectorstack{+8.2\\ \pm 6.2}} & {\tiny\Vectorstack{+12\\ \pm 13}} & {\tiny\Vectorstack{-3.3\\ \pm 9.6}} & {\tiny\Vectorstack{+10\\ \pm 7.1}} &  & {\tiny\Vectorstack{+6.6\\ \pm 7.2}} & {\tiny\Vectorstack{-5.5\\ \pm 6.6}} \\[2pt]
\textbf{\textcolor{t_darkgreen}{2-WL-GNN*}} ($\mathrm{SAM}$)&{\tiny\Vectorstack{+0.9\\ \pm 2.5}} & {\tiny\Vectorstack{-0.9\\ \pm 1.6}} & {\tiny\Vectorstack{+1.4\\ \pm 5.0}} & {\tiny\Vectorstack{+1.7\\ \pm 3.3}} & {\tiny\Vectorstack{+5.0\\ \pm 11}} & {\tiny\Vectorstack{-9.9\\ \pm 7.4}} & {\tiny\Vectorstack{+3.8\\ \pm 2.5}} & {\tiny\Vectorstack{-6.6\\ \pm 7.2}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-12\\ \pm 2.6}}} \\[2pt]
\textbf{2-WL-GNN\phantom{*}} ($\mathrm{SAM}$)&\cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+11\\ \pm 2.8}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+13\\ \pm 3.6}}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+14\\ \pm 3.1}}} & {\tiny\Vectorstack{+17\\ \pm 11}} & {\tiny\Vectorstack{+2.2\\ \pm 7.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 3.3}}} & {\tiny\Vectorstack{+5.5\\ \pm 6.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+12\\ \pm 2.6}}} & 
\end{tabular}}