### 2020-04-13 thesis/content/abstract.tex
improved intro and conclusion

%!TEX root = ../main.tex
%
\pdfbookmark[0]{Abstract}{Abstract}
\chapter*{Abstract}%
\label{sec:abstract}
\vspace*{-10mm}

This thesis describes the research field of \acl{gcr} from the perspective of the \ac{lta} problem.
It formally characterizes a selection of state-of-the-art \textit{graph kernels} and \acp{gnn} as instances of \ac{lta}.
Those characterizations are shown to be limited by the way in which they decompose graphs.
To overcome this limitation, an avenue for a more ``\acs{lta}-like'' \ac{gnn} is provided in form of so-called \textit{learned edge filters}.
To realize edge filters, the novel \textit{2-\acs*{wl}-\acs*{gnn}} model is proposed;
it is inspired by the two-dimensional \ac{wl} algorithm and proven to be strictly more expressive than existing \ac{gnn} approaches which are bounded by the more restrictive one-dimensional \ac{wl} algorithm.