### 2020-09-02 thesis/data/diffs/imdb.tex
added epoch timing evaluation infrastructure

% This file was generated by the LTAG results postprocessor. Do not edit manually.
{\setlength\tabcolsep{2.5pt}\setlength{\extrarowheight}{2pt}%
\begin{tabular}{lcccccccccc}
& \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{ST}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{WL\textsubscript{SP}} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-LWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{2-GWL} ($T=3$)} & \rotatebox[origin=l]{90}{\textbf{Baseline} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{GIN} ($\mathrm{sum}$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-GNN} ($\wmean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\mean$)} & \rotatebox[origin=l]{90}{\textbf{2-WL-GNN} ($\wmean$)} \\
\textbf{WL\textsubscript{ST}} ($T=3$)& & {\tiny\Vectorstack{-1.5\\ \pm 4.3}} & {\tiny\Vectorstack{+0.7\\ \pm 2.0}} & {\tiny\Vectorstack{+2.5\\ \pm 3.4}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 4.9}}} & {\tiny\Vectorstack{+6.1\\ \pm 6.4}} & {\tiny\Vectorstack{+1.5\\ \pm 2.6}} & {\tiny\Vectorstack{+2.0\\ \pm 2.7}} & {\tiny\Vectorstack{+1.7\\ \pm 3.7}} & {\tiny\Vectorstack{+1.8\\ \pm 3.8}} \\[2pt]
\textbf{WL\textsubscript{SP}} ($T=3$)&{\tiny\Vectorstack{+1.5\\ \pm 4.3}} &  & {\tiny\Vectorstack{+2.2\\ \pm 5.0}} & {\tiny\Vectorstack{+4.0\\ \pm 5.2}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+24\\ \pm 5.9}}} & {\tiny\Vectorstack{+7.6\\ \pm 7.4}} & {\tiny\Vectorstack{+3.0\\ \pm 5.2}} & {\tiny\Vectorstack{+3.5\\ \pm 5.7}} & {\tiny\Vectorstack{+3.2\\ \pm 6.4}} & {\tiny\Vectorstack{+3.3\\ \pm 6.7}} \\[2pt]
\textbf{2-LWL} ($T=3$)&{\tiny\Vectorstack{-0.7\\ \pm 2.0}} & {\tiny\Vectorstack{-2.2\\ \pm 5.0}} &  & {\tiny\Vectorstack{+1.8\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+22\\ \pm 5.7}}} & {\tiny\Vectorstack{+5.4\\ \pm 7.2}} & {\tiny\Vectorstack{+0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+1.1\\ \pm 3.7}} \\[2pt]
\textbf{2-GWL} ($T=3$)&{\tiny\Vectorstack{-2.5\\ \pm 3.4}} & {\tiny\Vectorstack{-4.0\\ \pm 5.2}} & {\tiny\Vectorstack{-1.8\\ \pm 2.6}} &  & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+3.6\\ \pm 7.1}} & {\tiny\Vectorstack{-1.0\\ \pm 2.6}} & {\tiny\Vectorstack{-0.5\\ \pm 4.1}} & {\tiny\Vectorstack{-0.8\\ \pm 3.1}} & {\tiny\Vectorstack{-0.7\\ \pm 4.3}} \\[2pt]
\textbf{Baseline} ($\mathrm{sum}$)&\cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-22\\ \pm 4.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-24\\ \pm 5.9}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 5.7}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} &  & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-16\\ \pm 6.3}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.0}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 5.6}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-21\\ \pm 6.4}}} & \cellcolor{t_red!25}\textcolor{t_darkred}{{\tiny\Vectorstack{-20\\ \pm 6.9}}} \\[2pt]
\textbf{GIN} ($\mathrm{sum}$)&{\tiny\Vectorstack{-6.1\\ \pm 6.4}} & {\tiny\Vectorstack{-7.6\\ \pm 7.4}} & {\tiny\Vectorstack{-5.4\\ \pm 7.2}} & {\tiny\Vectorstack{-3.6\\ \pm 7.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+16\\ \pm 6.3}}} &  & {\tiny\Vectorstack{-4.6\\ \pm 7.5}} & {\tiny\Vectorstack{-4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-4.3\\ \pm 8.4}} \\[2pt]
\textbf{2-GNN} ($\mean$)&{\tiny\Vectorstack{-1.5\\ \pm 2.6}} & {\tiny\Vectorstack{-3.0\\ \pm 5.2}} & {\tiny\Vectorstack{-0.8\\ \pm 2.0}} & {\tiny\Vectorstack{+1.0\\ \pm 2.6}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.0}}} & {\tiny\Vectorstack{+4.6\\ \pm 7.5}} &  & {\tiny\Vectorstack{+0.5\\ \pm 2.8}} & {\tiny\Vectorstack{+0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.2\\ \pm 2.4}} \\[2pt]
\textbf{2-GNN} ($\wmean$)&{\tiny\Vectorstack{-2.0\\ \pm 2.7}} & {\tiny\Vectorstack{-3.5\\ \pm 5.7}} & {\tiny\Vectorstack{-1.3\\ \pm 3.2}} & {\tiny\Vectorstack{+0.5\\ \pm 4.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 5.6}}} & {\tiny\Vectorstack{+4.1\\ \pm 7.1}} & {\tiny\Vectorstack{-0.5\\ \pm 2.8}} &  & {\tiny\Vectorstack{-0.3\\ \pm 3.5}} & {\tiny\Vectorstack{-0.2\\ \pm 2.7}} \\[2pt]
\textbf{2-WL-GNN} ($\mean$)&{\tiny\Vectorstack{-1.7\\ \pm 3.7}} & {\tiny\Vectorstack{-3.2\\ \pm 6.4}} & {\tiny\Vectorstack{-1.0\\ \pm 3.5}} & {\tiny\Vectorstack{+0.8\\ \pm 3.1}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+21\\ \pm 6.4}}} & {\tiny\Vectorstack{+4.4\\ \pm 7.9}} & {\tiny\Vectorstack{-0.2\\ \pm 2.8}} & {\tiny\Vectorstack{+0.3\\ \pm 3.5}} &  & {\tiny\Vectorstack{+0.1\\ \pm 3.4}} \\[2pt]
\textbf{2-WL-GNN} ($\wmean$)&{\tiny\Vectorstack{-1.8\\ \pm 3.8}} & {\tiny\Vectorstack{-3.3\\ \pm 6.7}} & {\tiny\Vectorstack{-1.1\\ \pm 3.7}} & {\tiny\Vectorstack{+0.7\\ \pm 4.3}} & \cellcolor{t_green!25}\textcolor{t_darkgreen}{{\tiny\Vectorstack{+20\\ \pm 6.9}}} & {\tiny\Vectorstack{+4.3\\ \pm 8.4}} & {\tiny\Vectorstack{-0.2\\ \pm 2.4}} & {\tiny\Vectorstack{+0.2\\ \pm 2.7}} & {\tiny\Vectorstack{-0.1\\ \pm 3.4}} & 
\end{tabular}}